---
title: "Covid91 vaccine study Final2023"
author: "Ambar Arreola"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

#methods
In this project we will effectiveness confidence interval graphs, numerical 

## Males

### Barplot

```{r}
dd2 <- male %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot illustrates the correlation between male participants and their susceptibility to contracting COVID-19 concerning the administration of the vaccine versus a placebo. The data reveals that approximately 70% of males who received the placebo contracted COVID-19, whereas those who received the vaccine had a lower contraction rate of around 30%.The OK category shows us that the people who did not recieve the vaccine or the placebo were both at around 50%. One of the plots shows us the count of males  
while the other shows the percentage.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=male)
rowPerc(table1)
colPerc(table1)
```
The table provides data on the percentage of individuals infected and not infected with COVID-19 for two different treatment groups: placebo and vaccine.
For the "COVID-19" row:Among those who received the placebo, 69.57% were infected with COVID-19, while 30.43% were not infected.Among those who received the vaccine, 30.43% were infected with COVID-19, while 69.57% were not infected.
For the "OK" row:Among those who received the placebo, 50.48% were not infected (recovered or not exposed to the virus), while 49.52% were infected.Among those who received the vaccine, 49.52% were not infected (recovered or not exposed to the virus), while 50.48% were infected.The "Total" row indicates that the percentages add up to 100% for both treatment groups, as it represents the total distribution of individuals between being infected with COVID-19 and being not infected (either recovered or not exposed) for both the placebo and vaccine groups.


### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
malefish <- fisher.test(table1)
malefish
```

The given test is Pearson's Chi-squared test with Yates' continuity correction. It analyzes the relationship between two categorical variables, "treatment" and "infection status" (covid19 or OK), based on the observed counts. The test yields a chi-square statistic of 55.183, with 1 degree of freedom and an extremely low p-value of 1.098e-13, indicating a highly significant association between the variables. The expected counts under the null hypothesis are also provided. Additionally, Fisher's Exact Test is performed, yielding a similar result and indicating that the odds ratio is significantly different from 1 with a 95% confidence interval of 1.798 to 2.811.

### Effectiveness Confidence Interval
```{r}
effect <- malefish$p.value/(malefish$p.value +1)
effectivenessguy = malefish$estimate/(malefish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(malefish$conf.int[1]/(malefish$conf.int[1] +1))*100
hi=(malefish$conf.int[2]/(malefish$conf.int[2] +1))*100
options(digits=2)
```

#### Effectiveness Confidence Interval `r low` to `r hi`

Here is the efficacy rate for the COVID-19 vaccine among males. If the efficacy is above 50%, it indicates that the vaccine is effective. Based on the data above we are able to see that the odds ration is 69% so this shows us that the vaccine is effective in protecting males from covid 19.

## Females

### Barplot


```{r}
dd2 <- female %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot depicts the relationship between female participants and their vulnerability to contracting COVID-19 in relation to receiving either the vaccine or a placebo. The data reveals that the vaccine was notably more efficacious in females compared to the placebo. Specifically, approximately 70% of females who received the placebo were infected with COVID-19, whereas only 40% of females who were vaccinated contracted the virus. Moreover, the plot indicates that 50% of females, regardless of whether they received the vaccine or placebo, remained unaffected by COVID-19. Both of these plots convey the same information, with the only distinction being the representation of data—one displays percentages, while the other exhibits the actual count of individuals.

### Numerical results
```{r}

table1 <- xtabs(~infected + treatment, data=female)
rowPerc(table1)
colPerc(table1)


```

The plot summarizes the outcomes of a treatment, comparing the number or percentage of individuals who were infected with COVID-19 and those who remained unaffected (OK) after receiving either a placebo or a vaccine. For the placebo group, out of 100 individuals, 60 were infected with COVID-19, while 50 remained unaffected. In contrast, for the vaccine group, out of 100 individuals, only 40 were infected with COVID-19, and 50 remained unaffected. This indicates that the vaccine was more effective in reducing the number of COVID-19 infections compared to the placebo.

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
femalefish <- fisher.test(table1)
femalefish
```

The plot presents the results of Pearson's Chi-squared test with Yates' continuity correction, comparing the observed counts of COVID-19 infections and non-infections (OK) in two treatment groups: placebo and vaccine. The Chi-square statistic is 13, with 1 degree of freedom, and a p-value of 3e-04, indicating a significant association between treatment and COVID-19 infection. Additionally, Fisher's Exact Test for Count Data was performed, yielding a p-value of 2e-04, and an estimated odds ratio of 1.5, suggesting a statistically significant difference between the two treatment groups regarding their susceptibility to COVID-19. The confidence interval for the odds ratio ranges from 1.2 to 1.9.

### Effectiveness confidence interval

```{r}
effect <- femalefish$p.value/(femalefish$p.value +1)
effectivenessfemale = femalefish$estimate/(femalefish$estimate +1)*100
print(effectivenessfemale)
```

```{r}
low=(femalefish$conf.int[1]/(femalefish$conf.int[1] +1))*100
hi=(femalefish$conf.int[2]/(femalefish$conf.int[2] +1))*100
options(digits=2)
```



#### Effectiveness Confidence Interval `r low` to `r hi`

The data provided reveals the efficacy rate of the COVID-19 vaccine in females. With an odds ratio of 60%, it indicates that the vaccine is effective in protecting females from COVID-19. Since the efficacy rate is above the 50% threshold, it confirms the vaccine's effectiveness. The confidence interval, spanning from 54.54% to 65.07%, further supports this conclusion, reinforcing the vaccine's ability to provide protection against COVID-19 in females.



## LGBTQ

### Barplot

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot illustrates the correlation between LGBTQ participants and their susceptibility to contracting COVID-19 in relation to receiving either the vaccine or a placebo. The data indicates that the vaccine's effectiveness was lower among the LGBTQ community compared to the placebo group. Specifically, among the LGBTQ participants who received the placebo, approximately 15% contracted COVID-19, while among those who received the vaccine, approximately 85% contracted the virus. Additionally, the plot shows that 50% of LGBTQ individuals, regardless of whether they received the vaccine or placebo, remained unaffected by COVID-19. Both plots convey the same information, with one displaying percentages and the other showing the actual count of individuals.

### Numerical results



```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

 Based on the given results, it appears to be a comparison of the effectiveness of a placebo treatment and a vaccine for COVID-19. The results show the number of individuals infected and not infected (OK) in each treatment group (placebo and vaccine).The "infected" row represents the number of individuals who contracted COVID-19 after receiving either the placebo or vaccine.The "placebo" column shows the number of infected individuals who received the placebo treatment.The "vaccine" column shows the number of infected individuals who received the COVID-19 vaccine.

    
  The second table shows the percentage of infected individuals in each treatment group, in relation to the total number of individuals in that group.For the infected row 0.66% of the indiviuals in the placebo group contracted covid 19.3.5% of the individuals in the vaccine group contracted covid 19.For the 99.34 of the individuals in the placebo group  and 96.5% in the vaccine group did not contract covid 19.In conclusion, the results indicate that the COVID-19 vaccine was significantly more effective than the placebo in preventing COVID-19 infections. The vaccine group had a lower percentage of infected individuals compared to the placebo group. Specifically, the infection rate was 3.5% in the vaccine group compared to 0.66% in the placebo group.

### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```

The table presents the results of a statistical analysis known as Pearson's Chi-squared test with Yates' continuity correction. It aims to investigate the relationship between two categorical variables: "infected" (with categories "covid19" and "OK") and "treatment" (with categories "placebo" and "vaccine") in the context of covid19 infection.The Chi-squared test yielded a significant Chi-square statistic of 29 with 1 degree of freedom and an extremely low p-value of 7e-08. This p-value suggests a highly significant association between "infected" and "treatment," indicating that the variables are not independent.The table contains observed counts for each combination of "infected" and "treatment," along with expected counts under the null hypothesis of independence. The contributions to the Chi-square statistic for each combination are also provided.Additionally, Fisher's Exact Test for Count Data was performed on the same dataset, yielding a p-value of 1e-08, further supporting the significant association between the variables.

In conclusion, the table demonstrates a strong and highly significant relationship between covid19 infection status and the type of treatment received (placebo or vaccine). The results provide robust evidence that the variables are related and not independent, suggesting that the treatment may influence the infection status in the context of covid19.



## Druggies


# Overall Results and Conclusions