---
title: "Covid91 vaccine study Final2023"
author: "Ambar Arreola"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

#methods
In this project we will effectiveness confidence interval graphs, numerical 

## Males

### Barplot

```{r}
dd2 <- male %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot illustrates the correlation between male participants and their susceptibility to contracting COVID-19 concerning the administration of the vaccine versus a placebo. The data reveals that approximately 70% of males who received the placebo contracted COVID-19, whereas those who received the vaccine had a lower contraction rate of around 30%.The OK category shows us that the people who did not recieve the vaccine or the placebo were both at around 50%. One of the plots shows us the count of males  
while the other shows the percentage.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=male)
rowPerc(table1)
colPerc(table1)
```
The table provides data on the percentage of individuals infected and not infected with COVID-19 for two different treatment groups: placebo and vaccine.
For the "COVID-19" row:Among those who received the placebo, 69.57% were infected with COVID-19, while 30.43% were not infected.Among those who received the vaccine, 30.43% were infected with COVID-19, while 69.57% were not infected.
For the "OK" row:Among those who received the placebo, 50.48% were not infected (recovered or not exposed to the virus), while 49.52% were infected.Among those who received the vaccine, 49.52% were not infected (recovered or not exposed to the virus), while 50.48% were infected.The "Total" row indicates that the percentages add up to 100% for both treatment groups, as it represents the total distribution of individuals between being infected with COVID-19 and being not infected (either recovered or not exposed) for both the placebo and vaccine groups.


### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
malefish <- fisher.test(table1)
malefish
```

The given test is Pearson's Chi-squared test with Yates' continuity correction. It analyzes the relationship between two categorical variables, "treatment" and "infection status" (covid19 or OK), based on the observed counts. The test yields a chi-square statistic of 55.183, with 1 degree of freedom and an extremely low p-value of 1.098e-13, indicating a highly significant association between the variables. The expected counts under the null hypothesis are also provided. Additionally, Fisher's Exact Test is performed, yielding a similar result and indicating that the odds ratio is significantly different from 1 with a 95% confidence interval of 1.798 to 2.811.

### Effectiveness Confidence Interval
```{r}
effect <- malefish$p.value/(malefish$p.value +1)
effectivenessguy = malefish$estimate/(malefish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(malefish$conf.int[1]/(malefish$conf.int[1] +1))*100
hi=(malefish$conf.int[2]/(malefish$conf.int[2] +1))*100
options(digits=2)
```

#### Effectiveness Confidence Interval `r low` to `r hi`

Here is the efficacy rate for the COVID-19 vaccine among males. If the efficacy is above 50%, it indicates that the vaccine is effective. Based on the data above we are able to see that the odds ration is 69% so this shows us that the vaccine is effective in protecting males from covid 19.

## Females

### Barplot


```{r}
dd2 <- female %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot depicts the relationship between female participants and their vulnerability to contracting COVID-19 in relation to receiving either the vaccine or a placebo. The data reveals that the vaccine was notably more efficacious in females compared to the placebo. Specifically, approximately 70% of females who received the placebo were infected with COVID-19, whereas only 40% of females who were vaccinated contracted the virus. Moreover, the plot indicates that 50% of females, regardless of whether they received the vaccine or placebo, remained unaffected by COVID-19. Both of these plots convey the same information, with the only distinction being the representation of data—one displays percentages, while the other exhibits the actual count of individuals.




## LGBTQ



## Druggies


# Overall Results and Conclusions